My wonderful paper

### 2024-11-15 _manuscript/_tex/index.tex
methodology

In real-world applications, it is rare to create a set of unit tests can
fully guarantee expected results. On one hand, there is the cost of
effort involved in manually developing all these tests; on the other,
there is the inherent complexity of the problem. (can we - if we're
thriving for detecting 95\% of the cause?). However, the quality of unit
tests can be evaluated by simulated data. One set of tests is considered
better than another if a small set of tests can reliably detect
unexpected outcomes, which brings two criteria in the evaluation metric:
accuracy and parsimony.

Accuracy refers to a set of tests' ability to accurately detect
unexpected outcomes while minimizing false positives and false
negatives. A false positive can indicate (caution or skepticism on
checking the data), whereas a false negative suggests the tests may lack
sensitivity to unexpected outcomes. Since both plan and outcome
expectations can be framed as unit tests with binary outcomes, one
approach is to predict the outcome expectation based on the plan
expectation. Performance of the tests can then be evaluated using
precision and recall metrics from the confusion matrix:

@@

\begin{itemize}
\item
  explain mutual information
\item
  explain logic regression
\item
  explain combining precision, recall, and independence together through
  ``means''
\item
  A logic regression (ref) is used to model the relationship between the
  plan and outcome expectations. (justify the use of Logic regression)
\end{itemize}

@@

\begin{itemize}
\item
  provide interpretation at different scenarios:

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi})}
  \tightlist
  \item
    one test is flagged, the prediction is as expected:
  \item
    multiple tests are flagged, the prediction is unexpected,
  \item
    no test is flagged, the prediction is unexpected,
  \item
    no test is flagged, the prediction is as expected
  \end{enumerate}
\end{itemize}

>>>>

In real-world applications, it is rare to create a set of unit tests can
fully guarantee expected results. On one hand, there is the cost of
effort involved in manually developing all these tests; on the other,
there is the inherent complexity of the problem. (can we - if we're
thriving for detecting 95\% of the cause?). However, the quality of unit
tests can be evaluated by simulated data. One set of tests is considered
better than another if a small set of tests can reliably detect
unexpected outcomes, which brings two criteria in the evaluation metric:
accuracy and parsimony. Figure~\ref{fig-metric-calc} illustrates the
workflow for calculating the metrics.

\phantomsection\label{cell-fig-metric-calc}
\begin{figure}[H]

\includegraphics[width=6.73in,height=\textheight]{figures/metric-calc.png}

\caption{\label{fig-metric-calc}this is the cap}

Accuracy refers to a set of tests' ability to accurately detect
unexpected outcomes while minimizing false positives and false
negatives. A false positive can indicate (caution or skepticism on
checking the data), whereas a false negative suggests the tests may lack
sensitivity to unexpected outcomes. To model the relationship between
the plan and outcome expectation (binary-binary), a logic regression
model is used \citep{ruczinski_logic_2003}. Originally developed for SNP
microarray data, logic regression constructs Boolean combinations of
binary variables to solve regression problems {[}more introduction on
logic regression{]}. Compared to other tree-based methods or machine
learning methods for binary-binary prediction, logic regression produces
Boolean combinations, or meta rules, that combines unit tests to solve
the prediction problem. {[}need rewrite here{]}. The performance of the
tests can be evaluated using precision and recall metrics derived from
the confusion matrix of the logic regression prediction

@@

To quantify redundancy, the concept of mutual information is used.
Mutual information \(I(x, y)\) measures the amount of information shared
between two random variables and is defined as the KL-distance
\(D(p \parallel q)\) between the joint distribution of the two variables
and the product of the marginal distributions:

\[I(x,y) = D\big(p(x,y) \parallel p(x)p(y)\big) = \sum_x \sum_y p(x,y) \log \frac{p(x,y)}{p(x)p(y)}\]

This concept extends naturally to multiple variables through total
correlation {[}ref{]}, \(C(X_1, X_2, \cdots, X_n)\), which captures
redundancy across a set of \(n\) variables:

\[C(X_1, X_2, \cdots, X_n) = \sum_{x_1} \sum_{x_2} \cdots \sum_{x_n} p(x_1, x_2, \cdots, x_n) \log \frac{p(x_1, x_2, \cdots, x_n)}{p(x_1)p(x_2) \cdots p(x_n)}\]

A high mutual information value indicates redundancy among the tests,
while a low value suggests that the tests are independent and provide
unique information to diagnose the unexpected outcome. To standardize
this measure, the total correlation \emph{per observation} is
calculated, and an independence score, ranging between 0 and 1, is
defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric,
various mathematical means, such as arithmetic mean, harmonic mean, and
quadratic mean, can be used. The differences among these means are
minimal when the three metrics are similar. However, as the differences
among the metrics increases, the harmonic mean tends to produce the
smallest overall score, as it penalizes low values, while the quadratic
mean tends to produce the largest score by rewarding higher values more.
For simple interpretation of the score, the arithmetic mean is
preferred, while in applications where the difference between precision,
recall, and independence need to be penalized or rewarded more, the
harmonic and quadratic mean should be considered.

@@

In the step count example, we can use the logic regression model to
evaluate the quality of the unit tests. The logic regression model is
fitted to the three unit tests (test1, test2, test3) and the outcome
expectation (unexpect) as the response variable. The model is then used
to predict the outcome expectation based on the unit tests. The
prediction is then compared to the actual outcome expectation to
calculate the precision and recall of the tests. The independence of the
tests is also calculated to assess the redundancy of the tests.
Figure~\ref{fig-logic-reg} presents the suggested logic regression model
as a combination of test 1 and test 3 with an OR operator.

Table~\ref{tbl-logic-reg} presents the calculated precision, recall, and
independence for the three individual tests and the combined test rule
(test1 OR test3) from the logic regression. The harmonic and arithmetic
means are included to evaluate the quality of the unit tests in
diagnosing unexpected step counts. The results show that the combined
test rule (test1 OR test3) has the highest precision, recall, and
independence, suggesting that it is the most effective test for
diagnosing unexpected step counts. We also include the metric calculated
from fitting a regression tree model to the data to compare the
performance of the logic regression model. The regression tree produces
a similar model of first split on test1 and then split on test3, and
results in the same accuracy and overall score as the logic regression
model.

\phantomsection\label{cell-fig-logic-reg}
\begin{figure}[H]

\includegraphics{index_files/figure-pdf/fig-logic-reg-1.pdf}

\caption{\label{fig-logic-reg}Logic regression model fitted to the three
unit tests (test1, test2, test3) and the outcome expectation (unexpect)
as the response variable. The model suggests using an OR rule to combine
test1 and test3 to predict the outcome expectation.}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2424}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1515}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1061}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1970}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1364}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}

\caption{\label{tbl-logic-reg}Accuracy (precision and recall) and
parsimony (independence) metrics for each individual unit test and for
the combined test rule (test1 OR test3) derived from the logic
regression model. The harmonic and arithmetic means of the three metrics
are included to evaluate the quality of the unit tests in diagnosing
unexpected step counts (more than five days with fewer than 8,000
steps).}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
tests
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
precision
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
independence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
harmonic
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
arithmetic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
test1 & 0.48 & 0.96 & 1 & 0.73 & 0.82 \\
test2 & 0.21 & 1.00 & 1 & 0.45 & 0.74 \\
test3 & 0.59 & 0.80 & 1 & 0.76 & 0.80 \\
test1 OR test3 & 0.82 & 0.84 & 1 & 0.88 & 0.89 \\
regression tree & 0.82 & 0.84 & 1 & 0.88 & 0.89 \\

\end{longtable}

### 2024-11-15 index.qmd
methodology

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(patchwork)
library(MASS)
library(ggh4x)
pm10 <- read_csv(here::here("data", "pm10.csv")) |> 
  filter(!is.na(pm10)) |>    
  mutate(season = as.factor(season))  
```

@@

In real-world applications, it is rare to create a set of unit tests can fully guarantee expected results. On one hand, there is the cost of effort involved in manually developing all these tests; on the other, there is the inherent complexity of the problem. (can we - if we're thriving for detecting 95% of the cause?). However, the quality of unit tests can be evaluated by simulated data. One set of tests is considered better than another if a small set of tests can reliably detect unexpected outcomes, which brings two criteria in the evaluation metric: accuracy and parsimony.

Accuracy refers to a set of tests' ability to accurately detect unexpected outcomes while minimizing false positives and false negatives. A false positive can indicate (caution or skepticism on checking the data), whereas a false negative suggests the tests may lack sensitivity to unexpected outcomes. Since both plan and outcome expectations can be framed as unit tests with binary outcomes, one approach is to predict the outcome expectation based on the plan expectation. Performance of the tests can then be evaluated using precision and recall metrics from the confusion matrix:

@@

- explain mutual information
- explain logic regression 
- explain combining precision, recall, and independence together through "means"

* A logic regression (ref) is used to model the relationship between the plan and outcome expectations. (justify the use of Logic regression)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(patchwork)
library(MASS)
library(ggh4x)
library(adtoolbox)
library(broom)
library(knitr)
library(rpart)
library(infotheo)
pm10 <- read_csv(here::here("data", "pm10.csv")) |> 
  filter(!is.na(pm10)) |>    
  mutate(season = as.factor(season))  
```

@@

In real-world applications, it is rare to create a set of unit tests can fully guarantee expected results. On one hand, there is the cost of effort involved in manually developing all these tests; on the other, there is the inherent complexity of the problem. (can we - if we're thriving for detecting 95% of the cause?). However, the quality of unit tests can be evaluated by simulated data. One set of tests is considered better than another if a small set of tests can reliably detect unexpected outcomes, which brings two criteria in the evaluation metric: accuracy and parsimony. @fig-metric-calc illustrates the workflow for calculating the metrics.

```{r}
#| label: fig-metric-calc
#| fig-cap: "this is the cap"
knitr::include_graphics(here::here("figures", "metric-calc.png"))
```

Accuracy refers to a set of tests' ability to accurately detect unexpected outcomes while minimizing false positives and false negatives. A false positive can indicate (caution or skepticism on checking the data), whereas a false negative suggests the tests may lack sensitivity to unexpected outcomes. To model the relationship between the plan and outcome expectation (binary-binary), a logic regression model is used [@ruczinski_logic_2003]. Originally developed for SNP microarray data, logic regression constructs Boolean combinations of binary variables to solve regression problems [more introduction on logic regression]. Compared to other tree-based methods or machine learning methods for binary-binary prediction, logic regression produces Boolean combinations, or meta rules, that combines unit tests to solve the prediction problem. [need rewrite here]. The performance of the tests can be evaluated using precision and recall metrics derived from the confusion matrix of the logic regression prediction

@@

To quantify redundancy, the concept of mutual information is used. Mutual information $I(x, y)$ measures the amount of information shared between two random variables and is defined as the KL-distance $D(p \parallel q)$ between the joint distribution of the two variables and the product of the marginal distributions:

$$I(x,y) = D\big(p(x,y) \parallel p(x)p(y)\big) = \sum_x \sum_y p(x,y) \log \frac{p(x,y)}{p(x)p(y)}$$

This concept extends naturally to multiple variables through total correlation [ref], $C(X_1, X_2, \cdots, X_n)$, which captures redundancy across a set of $n$ variables:

$$C(X_1, X_2, \cdots, X_n) = \sum_{x_1} \sum_{x_2} \cdots \sum_{x_n} p(x_1, x_2, \cdots, x_n) \log \frac{p(x_1, x_2, \cdots, x_n)}{p(x_1)p(x_2) \cdots p(x_n)}$$

A high mutual information value indicates redundancy among the tests, while a low value suggests that the tests are independent and provide unique information to diagnose the unexpected outcome. To standardize this measure, the total correlation *per observation* is calculated, and an independence score, ranging between 0 and 1, is defined as 1 - mutual information.

To combine precision, recall, and independence into a single metric, various mathematical means, such as arithmetic mean, harmonic mean, and quadratic mean, can be used. The differences among these means are minimal when the three metrics are similar. However, as the differences among the metrics increases, the harmonic mean tends to produce the smallest overall score, as it penalizes low values, while the quadratic mean tends to produce the largest score by rewarding higher values more. For simple interpretation of the score, the arithmetic mean is preferred, while in applications where the difference between precision, recall, and independence need to be penalized or rewarded more, the harmonic and quadratic mean should be considered.

@@

```{r}
#| label: fig-logic-reg
#| fig-cap: Logic regression model fitted to the three unit tests (test1, test2, test3) and the outcome expectation (unexpect) as the response variable. The model suggests using an OR rule to combine test1 and test3 to predict the outcome expectation.
fit <- step_count |> fit_logic_reg(unexpect, test1:test3, seed = 1, nleaves = 4)
plot(fit)
```

```{r}
#| label: tbl-logic-reg
list(tibble(.fitted = step_count$test1, unexpect = step_count$unexpect),
     tibble(.fitted = step_count$test2, unexpect = step_count$unexpect),
     tibble(.fitted = step_count$test3, unexpect = step_count$unexpect)
     ) |>
  map_dfr(~.x |> calc_miscla_rate(unexpect, .fitted) |> 
            mutate(independence = 1) |> 
            calc_metrics(metrics = c("harmonic", "arithmetic")), .id = "tests") |> 
  mutate(tests = c("test1", "test2", "test3")) |>
  bind_rows(augment(fit) |> 
              calc_miscla_rate(unexpect, .fitted) |> 
              calc_independence() |> 
              calc_metrics(metrics = c("harmonic", "arithmetic")) |> 
              mutate(tests = "test1 OR test3") |> 
              dplyr::select(-overlapping)) |>
  kable(digits = 3, 
        caption = "Accuracy (precision and recall) and parsimony (independence) metrics for each individual unit test and for the combined test rule (test1 OR test3) derived from the logic regression model. The harmonic and arithmetic means of the three metrics are included to evaluate the quality of the unit tests in  diagnosing unexpected step counts (more than five days with fewer than 8,000 steps).")
```
